\documentclass[a4paper,11pt]{article}
\usepackage[utf8]{inputenc} % Allows UTF-8 input
\usepackage[T1]{fontenc}    % Use 8-bit T1 fonts
\usepackage{geometry}       % Adjusts margins
\geometry{margin=1in}
\usepackage{amsmath, amssymb} % Math packages
\usepackage{enumitem}       % Customizable lists
\usepackage{hyperref}       % For hyperlinks
\usepackage{titlesec}       % Section customization

% Customize section titles
\titleformat{\section}{\large\bfseries}{\thesection.}{1em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection.}{1em}{}

% Document information
\title{Mathematics Literature for Study}
\author{Your Name}
\date{\today}

\begin{document}
\maketitle

\section*{Introduction}
This document provides a curated list of books and resources for studying mathematics, ranging from foundational topics to advanced concepts. Each entry includes the book title, author(s), and a brief description of its significance.

\section{Foundational Mathematics}
\begin{itemize}[left=0pt]
    \item \textbf{Basic Mathematics}\\
    Author: Serge Lang\\
    A solid introduction to fundamental mathematical concepts, ideal for those seeking a strong base.
    
    \item \textbf{How to Prove It: A Structured Approach}\\
    Author: Daniel J. Velleman\\
    A clear and practical guide for learning mathematical proof techniques.
\end{itemize}

\section{Intermediate Topics}
\begin{itemize}[left=0pt]
    \item \textbf{Introduction to Real Analysis}\\
    Author: Bartle and Sherbert\\
    A rigorous exploration of real analysis, perfect for students transitioning to higher mathematics.
    
    \item \textbf{Linear Algebra Done Right}\\
    Author: Sheldon Axler\\
    A modern and conceptual approach to linear algebra without determinants.
\end{itemize}

\section{Advanced Topics}
\begin{itemize}[left=0pt]
    \item \textbf{Principles of Mathematical Analysis}\\
    Author: Walter Rudin\\
    A classic text covering real and complex analysis at an advanced undergraduate level.

    \item \textbf{Abstract Algebra}\\
    Author: David S. Dummit and Richard M. Foote\\
    A comprehensive and in-depth resource for understanding abstract algebraic structures.
\end{itemize}

\section{Specialized Mathematics}
\begin{itemize}[left=0pt]
    \item \textbf{Number Theory: An Introduction via the Distribution of Primes}\\
    Author: Benjamin Fine\\
    A gentle introduction to number theory with an emphasis on prime numbers.
    
    \item \textbf{Topology}\\
    Author: James Munkres\\
    A highly regarded resource for understanding the basics of topology.
\end{itemize}

\section*{Conclusion}
This list serves as a roadmap for mathematical studies, from foundational topics to more specialized areas. Explore these resources to build a solid mathematical foundation and develop advanced skills.

\end{document}
